---
title: "R Notebook Income Dataset"
output: html_notebook
---
## Readcsv file
```{r}
```{r}
library(readr)
df <- read_csv("C:/Users/Rushi/Downloads/income.csv")[,-c(1)]
head(df)
str(df)
```
### EDA of bivariate data

Calculation of correlation coefficient using **correlate()**
correlate() calculates the correlation coefficient of all combinations of df numerical variables as follows:
```{r}
library(dlookr)
describe(df)
correlate(df)
```
## EDA based on target variable
**Definition of target variable**
To perform EDA based on target variable, you need to create a target_by class object. target_by() creates a target_by class with an object inheriting data.frame or data.frame. target_by() is similar to group_by() in dplyr which creates grouped_df. The difference is that you specify only one variable.

The following is an example of specifying CrimeRate as target variable in Crime data.frame.:
```{r}
target <- target_by(df, happiness)
target
```
## EDA when target variable is categorical variable
**CASES WHERE PREDICTORS ARE NUMERIC VARIABLE**
> relate() shows the relationship between the target variable and the predictor. The following example shows the relationship between happiness and the target variable income. The predictor happiness is a numeric variable. In this case, the descriptive statistics are shown for each level of the target variable.

```{r}
# If the variable of interest is a numerical variable
dff <- relate(target, income)
dff
```
## Find summary of the linear model
```{r}
summary(lm(happiness~income,df))
```

## Plotting normality of target variable
```{r}
plot_normality(df, happiness)
```
```{r}
lmdist = lm(happiness~income, df) #Create the linear regression
plot(df$income,df$happiness, col = "yellow")+abline(lmdist) #Add a regression line
```
Find the best fitting for the target variable
```{r}
library(dplyr)
library(tidypredict)
model <- lm(happiness~income, df)
df %>% tidypredict_to_column(model, add_interval = TRUE) %>% head(10)
```
### **Inference -** We can observe happiness has high correlation

### Visualization of the correlation matrix using plot_correlate()
plot_correlate() visualizes the correlation matrix.
```{r}
plot_correlate(df)
```