---
title: "GGPlot on MidWest Dataset"
output: html_notebook
---

```{r}
library('readr')
df = read_csv('C:\\Users\\Rushi\\Desktop\\midwest.csv')
head(df,10)
```
```{r}
library(ggplot2)
ggplot(df, aes(x=area, y=poptotal)) + geom_point()
```
We got a basic scatterplot, where each point represents a county. 

Let's just add a smoothing layer using geom_smooth(method='lm'). 
Since the method is set as lm (short for linear model), it draws the line of best fit.
```{r}
g <- ggplot(midwest, aes(x=area, y=poptotal)) + geom_point() + geom_smooth(method="lm",se=FALSE) # set se=FALSE to turnoff confidence bands
plot(g)

```
You might have noticed that majority of points lie in the bottom of the chart which doesn't really look nice. So, let's change the Y-axis limits to focus on the lower half.

## **Adjusting the X and Y axis limits**
### Method 1: By deleting the points outside the range
> This will change the lines of best fit or smoothing lines as compared to the original data.
> This can be done by **xlim() and ylim()**. You can pass a numeric vector of length 2 (with max and min values) or just the max and min values itself.

```{r}
g <- ggplot(midwest, aes(x=area, y=poptotal)) + geom_point() + geom_smooth(method="lm") # set se=FALSE to turnoff confidence bands

# Delete the points outside the limits
g + xlim(c(0, 0.1)) + ylim(c(0, 1000000)) # deletes points

```
> In this case, the chart was not built from scratch but rather was built on top of g. This is because, 
the previous plot was stored as g, a ggplot object, which when called will reproduce the original 
plot. 
> Did you notice that the line of best fit became more horizontal compared to the original plot? This 
is because, when using xlim() and ylim(), the points outside the specified range are deleted and 
will not be considered while drawing the line of best fit

### Method 2: Zooming In
The other method is to change the X and Y axis limits by zooming in to the region of interest 
without deleting the points. This is done using **coord_cartesian()**.
```{r}
g <- ggplot(midwest, aes(x=area, y=poptotal)) + geom_point() + geom_smooth(method="lm") # set se=FALSE to turnoff confidence bands
# Zoom in without deleting the points outside the limits. 
# As a result, the line of best fit is the same as the original plot.
g1 <- g + coord_cartesian(xlim=c(0,0.1), ylim=c(0, 1000000)) # zooms in
plot(g1)

```
Since all points were considered, the line of best fit did not change.
## **How to Change the Title and Axis Labels**
> I have stored this as g1. Let's add the plot title and labels for X and Y axis. This can be done in one go using the labs() function with title, x and y arguments. Another option is to use the **ggtitle(), xlab() and ylab()**.

```{r}
# Add Title and Labels
g1 + labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population", x="Area", caption="Midwest Demographics", alt = 'midwest dataset')
```
```{r}
# or
g1+ggtitle("Area Vs Population", subtitle="From midwest dataset")+xlab("Area") + ylab("Population")
```
Full Plot Call
```{r}
# Full Plot call
library(ggplot2)
ggplot(midwest, aes(x=area, y=poptotal)) +
 geom_point() +
 geom_smooth(method="lm") +
 coord_cartesian(xlim=c(0,0.1), ylim=c(0, 1000000)) +
 labs(title="Area Vs Population", 
      subtitle="From midwest dataset",
      y="Population", 
      x="Area", 
      caption="Midwest Demographics")
```
## **How to Change the Color and Size of Points**
```{r}
ggplot(midwest, aes(x=area, y=poptotal)) +
 geom_point(col="pink", size=3) + # Set static color and size for points
 geom_smooth(method="lm", col="firebrick") + # change the color of line
 coord_cartesian(xlim=c(0, 0.1), ylim=c(0, 1000000)) +
 labs(title="Area Vs Population", 
      subtitle="From midwest dataset",
      y="Population", 
      x="Area", 
      caption="Midwest Demographics")


```
## **How to Change the Color To Reflect Categories in Another Column?**
```{r}
gg <- ggplot(midwest, aes(x=area, y=poptotal)) +
 geom_point(aes(col=state), size=3) + # Set color to vary based on state categories.
 geom_smooth(method="lm", col="firebrick", size=2) +
 coord_cartesian(xlim=c(0, 0.1), ylim=c(0, 1000000)) +
 labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population"
, x="Area", caption="Midwest Demographics")
plot(gg)
```

