---
title: "11/03/2022 DataFrame Practice"
output: html_notebook
---

### Create DataFrame with 3 columns name,age and gender
```{r}
df=data.frame(name=letters[1:5],age=runif(5,min=0,max=99),gender=c('M','F','F','M','F'))
df
```
### Extract details of name, age
```{r}
summary(df$name)
str(df$name)
```
```{r}
summary(df$age)
str(df$age)
```
### Print details of eldest person
```{r}
df[which.max(df$age),]
```
### Print details of 2nd column
```{r}
df[,2]
```
### Remove third, forth column
```{r}
df[,-4]
```
### Print str of df
```{r}
str(df)
```
### Number od columns in dataframe
```{r}
ncol(df)
```
### Number of rows in dataframe
```{r}
nrow(df)
```
### Dimension of DataFrame
```{r}
dim(df)
```
### Display column names
```{r}
names(df)
```
### Indexing through rows and columns
```{r}
df[2] #2nd column 
df[2:3,] # 2,3 row of all columns
df["name"] # name column
df[2:3,c("age")] # 2nd and 3rd row of age column
```
### Fetch multiple columns
```{r}
df[c("age",'gender')]
```
### Fetch multiple columns and can set column names with "column" attribute
```{r}
data.frame(df$name,df$age)
```
---
### Help on ChickWeight dataset
```{r}
?ChickWeight
```
### Extract features
```{r}
names(ChickWeight)
```
### Use view() to view dataset
```{r}
View(ChickWeight)
```
### head() | tail() | str() | summary() | frequency() | time() of dataset
```{r}
head(ChickWeight) #first 6 records
```
```{r}
tail(ChickWeight) # last 6 records
```
```{r}
str(ChickWeight) # str of dataset
```
```{r}
summary(ChickWeight) # summary like mean,med, mode etc. of dataset
```
```{r}
#It gives the no. of occurrences of every unique value in the variable.
frequency(ChickWeight)
```
```{r}
# time() creates the vector of times at which a time series was sampled.
time(ChickWeight$Time)
```
