---
title: "R Notebook"
output: html_notebook
---
### Read Uber Ride Data choosing a file
```{r}
rides=read.csv(file.choose())
```
### View the whole dataset in R
```{r}
View(rides)
```
### Get Dataset Info
```{r}
str(rides)
```
# Data Cleaning:
## While analyzing the first column of the division, I clearly saw that more work was needed, because I could find different values ​​referring to the same category. After that, I summarized the first 15 paragraphs out of 5.

### Checking categories in product_type column
```{r}
rides %>% group_by(ProductType) %>% summarise(value_counts = n())
```
### Categories Reclassification
```{r}
prodMap= c("UberX"="UberX","uberX"="UberX","uberX VIP"="UberX","VIP"="UberX","POOL"="Pool","POOL: MATCHED"="Pool","UberBLACK"="Black","uberx"="UberX","uberPOOL"="Pool","uberPOOL: MATCHED"="Pool","Pool: MATCHED"="Pool")
as.character(prodMap)# get values
as.character(names(prodMap)) #get keys
```
### New categories replacement
```{r}
column_to_replace=as.character(names(prodMap))
replacement_value=as.character(prodMap)
#rides <- as.data.frame(t(rides))
replace(rides$ProductType,replacedata,datatobereplaced)
data.frame(rides$ProductType)
rides['ProductType']
list(column_to_replace)
rides[rides$ProductType == list(column_to_replace)] <- list(replacement_value)
data.frame(rides$ProductType %>% list(column_to_replace))
```

