---
title: "EDA on crime dataset"
output: html_notebook
---
# dlookr Package
Introduce dlookr package for explore data to understand the data and find scenarios for performing the analysis.

The dlookr package makes these steps fast and easy:

**(+) dlookr increases synergy with dplyr. Particularly in data exploration and data wrangle, it increases the efficiency of the tidyverse package group.**

> ###### 1. Performs an data diagnosis or automatically generates a data diagnosis report.
> ###### 2. **Discover data in a variety of ways, and automatically generate EDA(exploratory data analysis) report.**
> ###### 3. Impute missing values and outliers, resolve skewed data, and categorize continuous variables into categorical variables. And generates an automated report to support it.

```{r}
library(readr)
df <- read_csv("C:/Users/Rushi/Downloads/Crime_R.csv")
head(df)
str(df)
```
### Use Skimr to Display Summary Statistics for the variables

This has same as above + missing and a histogram.  Also, it has some additional statistics of non-numerical values.
```{r}
library(skimr)
skim(df)
```
### Drop unecessary columns holding with col_name10 from dataset
```{r}
df <- df[, -c(15:27)] # delete columns 5 through 7
colnames(df)
```

##### The following is a list of the *EDA functions* included in the dlookr package.:

>##### 1. **describe()** provides descriptive statistics for numerical data.
>##### 2. **normality()** and plot_normality() perform normalization and visualization of numerical data.
>##### 3. **correlate()** and plot_correlate() calculate the correlation coefficient between two numerical data and provide visualization.
>##### 4. **target_by()** defines the target variable and relate() describes the relationship with the variables of interest corresponding to the target variable.
>##### 5. **plot.relate()** visualizes the relationship to the variable of interest corresponding to the destination variable.
>##### 6. **eda_report()** performs an exploratory data analysis and reports the results.

## **Univariate data EDA**
---
### Calculating descriptive statistics using describe()

 describe() computes descriptive statistics for numerical data. The descriptive statistics help determine the distribution of numerical variables.
The variables of the df object returned by describe() are as follows.

**n** : number of observations excluding missing values
**na** : number of missing values
**mean** : arithmetic average
**sd** : standard deviation
**se_mean** : standard error mean. sd/sqrt(n)
**IQR** : interquartile range (Q3-Q1)
**skewness** : skewness
**kurtosis** : kurtosis
**p25** : Q1. 25% percentile
**p50** : median. 50% percentile
**p75** : Q3. 75% percentile
**p01, p05, p10, p20, p30** : 1%, 5%, 20%, 30% percentiles
**p40, p60, p70, p80** : 40%, 60%, 70%, 80% percentiles
**p90, p95, p99, p100** : 90%, 95%, 99%, 100% percentiles
```{r}
library(dlookr)
describe(df)
```
### EDA of bivariate data

Calculation of correlation coefficient using **correlate()**
correlate() calculates the correlation coefficient of all combinations of df numerical variables as follows:
```{r}
correlate(df)
```
### **Inference -** We can observe ExpenditureYear0 had high correlation

### Visualization of the correlation matrix using plot_correlate()
plot_correlate() visualizes the correlation matrix.
```{r}
plot_correlate(df)
```
### The plot_correlate() function also supports the group_by() function syntax in the dplyr package.
```{r}
library('magrittr')
library('dplyr')
df %>%
  filter(MoreMales == '0') %>%
  group_by(Southern) %>%
  plot_correlate(CrimeRate)
```
## Visualization of normality of numerical variables using **plot_normality()**
plot_normality() visualizes the normality of numeric data.

The information visualized by plot_normality() is as follows.:

>1. Histogram of original data
>2. Q-Q plot of original data
>3. histogram of log transformed data
>4. Histogram of square root transformed data

In the data analysis process, it often encounters numerical data that follows the power-law distribution. Since the numerical data that follows the power-law distribution is converted into a normal distribution by performing the log or sqrt transformation, so draw a histogram of the log and sqrt transformed data.

**plot_normality()** can also specify several variables like normality() function.
```{r}
# Select columns by name
plot_normality(df, CrimeRate, Wage)
```
```{r}
# Select single column
plot_normality(df, CrimeRate)
```
## EDA based on target variable
**Definition of target variable**
To perform EDA based on target variable, you need to create a target_by class object. target_by() creates a target_by class with an object inheriting data.frame or data.frame. target_by() is similar to group_by() in dplyr which creates grouped_df. The difference is that you specify only one variable.

The following is an example of specifying CrimeRate as target variable in Crime data.frame.:
```{r}
target <- target_by(df, CrimeRate)
target
```
## EDA when target variable is categorical variable
**CASES WHERE PREDICTORS ARE NUMERIC VARIABLE**
> relate() shows the relationship between the target variable and the predictor. The following example shows the relationship between Education and the target variable CrimeRate. The predictor Education is a numeric variable. In this case, the descriptive statistics are shown for each level of the target variable.

```{r}
# If the variable of interest is a numerical variable
dff <- relate(target, Education)
dff
```

