---
title: "EDA on crime dataset"
output: html_notebook
---
# dlookr Package
Introduce dlookr package for explore data to understand the data and find scenarios for performing the analysis.

The dlookr package makes these steps fast and easy:

**(+) dlookr increases synergy with dplyr. Particularly in data exploration and data wrangle, it increases the efficiency of the tidyverse package group.**

> ###### 1. Performs an data diagnosis or automatically generates a data diagnosis report.
> ###### 2. **Discover data in a variety of ways, and automatically generate EDA(exploratory data analysis) report.**
> ###### 3. Impute missing values and outliers, resolve skewed data, and categorize continuous variables into categorical variables. And generates an automated report to support it.

```{r}
library(readr)
df <- read_csv("C:/Users/Rushi/Downloads/Crime_R.csv")
head(df)
str(df)
```
### Use Skimr to Display Summary Statistics for the variables

This has same as above + missing and a histogram.  Also, it has some additional statistics of non-numerical values.
```{r}
library(skimr)
skim(df)
```
### Drop unecessary columns holding with col_name10 from dataset
```{r}
df <- df[, -c(15:27)] # delete columns 5 through 7
colnames(df)
```

##### The following is a list of the *EDA functions* included in the dlookr package.:

>##### 1. **describe()** provides descriptive statistics for numerical data.
>##### 2. **normality()** and plot_normality() perform normalization and visualization of numerical data.
>##### 3. **correlate()** and plot_correlate() calculate the correlation coefficient between two numerical data and provide visualization.
>##### 4. **target_by()** defines the target variable and relate() describes the relationship with the variables of interest corresponding to the target variable.
>##### 5. **plot.relate()** visualizes the relationship to the variable of interest corresponding to the destination variable.
>##### 6. **eda_report()** performs an exploratory data analysis and reports the results.

## **Univariate data EDA**
---
### Calculating descriptive statistics using describe()

 describe() computes descriptive statistics for numerical data. The descriptive statistics help determine the distribution of numerical variables.
The variables of the df object returned by describe() are as follows.

**n** : number of observations excluding missing values
**na** : number of missing values
**mean** : arithmetic average
**sd** : standard deviation
**se_mean** : standard error mean. sd/sqrt(n)
**IQR** : interquartile range (Q3-Q1)
**skewness** : skewness
**kurtosis** : kurtosis
**p25** : Q1. 25% percentile
**p50** : median. 50% percentile
**p75** : Q3. 75% percentile
**p01, p05, p10, p20, p30** : 1%, 5%, 20%, 30% percentiles
**p40, p60, p70, p80** : 40%, 60%, 70%, 80% percentiles
**p90, p95, p99, p100** : 90%, 95%, 99%, 100% percentiles
```{r}
library(dlookr)
describe(df)
```

