---
title: "Car Evaluation Data Analysis"
output: html_notebook
---

## Problem : 
To analyze the given dataset “Car Evaluation Database” and build a ML model using ML concepts to check it’s accuracy.

## Contents :

1. Import the Dataset
2. Clean the Dataset
3. Plots on Dataset
4. Spliting the Dataset
5. Building the Model
6. Accuracy of Model

# Install Tidyverse, Caret, RPart for Recursive Partioning Decision Trees
```{r setup='false'}
library(tidyverse)
library(caret)
library(rpart)
```

# Load Dataset
```{r}
df=read.csv('C:\\Users\\Rushi\\Downloads\\CSVs\\car_evaluation.csv',stringsAsFactors = T)
df
```
## Check the data
```{r}
sum(is.na(df))
sum(duplicated(df))
```
## Naming the columns 
```{r}
colnames(df)=c("buying","maint","doors","persons","lug_boot","safety","class")
set.seed(777)
```

## Now lets see the dimensions,a glimpse and summary of the given dataset
```{r}
#Summary
library(dplyr)
#dimensions of the dataset
dim(df)
#simple glimpse of the dataset
glimpse(df)
#summary of each column
summary(df)
```
# EDA using esquisse package
```{r}
#To get the data
library(yarrr)

#To get the exploratory package
library(esquisse)

#To launch the UI and explore the pirates data
esquisse::esquisser(df, viewer = "browser")
```

```{r}
#Visualisation
library(ggplot2)
ggplot(df,aes(x=class,fill=lug_boot))+geom_histogram(stat="count")+labs(title="Class Vs Luggage boot",subtitle="Histogram",y="Frequency of Luggage boot",x="Class")
```
```{r}
ggplot(data = df,aes(fill=as.factor(doors),x=persons))+geom_density(alpha=0.3)
```
```{r}
ggplot(data = df,aes(fill=as.factor(maint),x=class))+geom_density(alpha=0.3)+facet_wrap(~class)
```
# Splitting the Dataset

### Since this is a large data, we split it as training and testing sets. Thus we can build a model using training set and use it to predict results on testing set.
```{r}
#Fitting the model
library(caTools)
```
```{r}
shuffle_index=sample(1:nrow(df))
df=df[shuffle_index,]
split=sample.split(df$class,SplitRatio = 0.8)
head(split)
```
```{r}
#Splitting the dataset into testing and training parts
training_set=subset(df,split==TRUE)
testing_set=subset(df,split==FALSE)

View(training_set)
View(testing_set)
```

# Building Model Using Machine Learning Algorithm: Decision Tree

### Fitting the model as Decision Tree : **Rpart(Recursive partitioning is a statistical method for multivariable analysis)** is a powerful machine learning library in R that is used for building classification and regression trees. This library implements recursive partitioning and is very easy to use
```{r}
#Decision Tree
library(rpart)
library(rpart.plot)

fit=rpart(formula = class ~ .,data=training_set,method = "class")
predict_unseen=predict(object = fit,newdata = testing_set,type = "class")
fit1=rpart(formula = class ~ (safety+buying+persons),data=training_set,method = "class")
fit2=rpart(formula = class ~ (safety+maint+persons),data=training_set,method = "class")
fit3=rpart(formula = class ~ (safety+buying+maint+persons),data=training_set,method = "class")
```
## Visualizing the Decision Trees
```{r}
rpart.plot(fit)
```
```{r}
rpart.plot(fit1)
```
# Building Model Using Machine Learning Algorithm: Logistic Regression
```{r}
#Logistic Regression
linear_regressor=glm(safety~.,family='quasibinomial',training_set)
summary(linear_regressor)
```

```{r}
library(inspectdf)
inspect_types(df) %>% show_plot()
```

```{r}
prob_predict=predict.glm(object=linear_regressor,newdata=testing_set,type = "response")
plot(training_set$safety,training_set$class)
```

# Prediction made for the model,
```{r}
#Confusion matrix
cm=table(testing_set$class,predict_unseen)
cm
```

## Lets find the accuracy and misclassification in the developed model,
```{r}
accuracy=(cm[1,1]+cm[2,2]+cm[3,3]+cm[4,4])/NROW(testing_set)
accuracy
```
```{r}
misclassification=1-accuracy
misclassification
```
```{r}
#Pie Chart showing the Class Distribution 
acc=cm[1,1]
good=cm[2,2]
unacc=cm[3,3]
vgood=cm[4,4]
library(plotrix)
pie3D(x=c(acc,good,unacc,vgood),radius=1,explode = 0.05,labels = c("Accountable","Good","Un-Accountable","Very Good"),main="Condition of class in testing_set",start=67)
```
# Conclusion

* All the variables are importance for customers in assessing whether the car is in acceptable or unacceptable range.
* Safety and Seating capacity are two main factors in rejecting the cars as unacceptable.
* Number of doors are the least important variable in deciding the class value of the car.
