---
title: "R for Beginners"
output:
  html_document:
    df_print: paged
---

Describes the packages

```{r}
demo()
```

To list the demos in all *available* packages.

```{r}
demo(package = .packages(all.available = TRUE))
```

Default working directory path

```{r}
getwd()
```

Set NEW working directory path

```{r}
setwd('C:/Users/Rushi/Downloads')
getwd()
```

Help on functions with ? or ?? 

```{r}
?mean()
??mean()
```

Startup tutorial in R documentation

```{r}
help.start()
```

Help on the package help(PACKAGE_NAME)

```{r}
help(plotmath)
```

Gives an executable example of mean function

```{r}
example(mean)
```

Lists all funcs containing atleast "mean()" string

```{r}
apropos("plot")
```

Assigning Different Values

```{r}
x="AMD"
y=4
z=4.4
print(x)
print(y)
print(z)
```

Expressions and assignments in R

```{r}
x=5
y=x+10
y
```

Listing all user-defined objects

```{r}
ls()
```

Printing - Auto/Explicit

```{r}
a=3
b=5
c=9
sum=a+b+c
sum # Auto Printing
print(sum) #Explicit printing
```

Remove Objects

```{r}
rm(x,y)
ls()
```

Creating vector;checking type, class, is.integer and converting to integer

```{r}
# Create Vector
x<-c(1,2,3,4,5,6,7,8,9,10)
# Check the class of object
class(x)
#Check type of object
typeof(x)
# Is the value integer or not!
is.integer(x)
# Convert any datatype to int
y=as.integer(x)
typeof(y)
```

Assigning Tuple to variable and accessing through index

```{r}
assign("AMD",c(4,44,444))
AMD
AMD[0:3]
```

Auto Type Casting

```{r}
# Vectors can store any 1 type, if we assign more than 1 type it converts to higher type
assign("BEC",c(4,44,444,'D'))
class(BEC)
BEC[0:4]
assign("ABC",c(4,44,4.4))
class(ABC)
ABC[0:3]
```

Vector Operations

```{r}
c(1,2,3)->x
x
x+1
x*10
1/x
```

Extending the vector

```{r}
y=c(x,c(7,8,9))
y
x
y
x+y
```

Task-Create 2 vectors and mul 2nd vector with 3 and combine and find length

```{r}
assign("a",c(1,2,3,4,5))
assign("b",c(6,7,8,9,10))
length(c(a,b*3))
```

Find the 5th power of any vector

```{r}
a^5
```

Sum of all values in a vector

```{r}
sum(a)
```

Range of values

```{r}
range(a)
```

Mean from vector

```{r}
mean(a)
```

Variance of vector

```{r}
var(a)
```

Ascending Sort

```{r}
sort(a)
```

Descending Sort

```{r}
sort(a,decreasing=TRUE)
```

Find Squareroot

```{r}
sqrt(a)
```

Sequences - Changing the order of parameters

```{r}
#Create sequence with steps
seq(3,100,3) # Display multiples of 3
```

```{r}
seq(50,)
```
```{r}
seq(50,by=1,length=100)
```
```{r}
seq(2, 20, by=.2)
```
```{r}
seq(2, 20, length=2)
```
```{r}
seq(2,20, length=5) 
```
```{r}
seq(2, by=1, length=5)
```
```{r}
seq(2, by=1, length=15)
```
```{r}
seq(2, by=2, length=15)
```
```{r}
seq(by=2, length=15,15)
```
```{r}
seq(by=2, length=15,to=15) 
```
```{r}
seq(by=2, length=15,from=15)
```
```{r}
seq(length=51, from=-5, by=.2)
```

Logical Operators

```{r}
assign("a",c(1,2,3,4,5,2))
assign("b",c(2,7,1,4,0,0))
a&b # And Operation
a|b # Or Operation
```
Check Missing Values i.e. NA/NaN in Data

```{r}
z=c(NA,1:5,NA,0/0)
z
is.na(z) # Returns True where NA(Missing value/s) is found
is.nan(z) # Returns True if NaN(Not a Number) value is found leading to infinite answer
```
Escape Sequence Printing

```{r}
str="Good Morn!\n This is your fav sender :)\\^.^/"
print(str) # It prints whatever is embedded with single/double quotes
writeLines(str) # It executes the escape sequences like \n,\,/...
```
Zipping values with/without separator

```{r}
paste(c("$"),1:10)
paste(0:3,1:3,sep=",")
paste(c("x","y","z"),1:10,sep=",")
```
Omit *Missing Values* and fetch rest values

```{r}
vector_with_na=c(NA,0:5,9:10,NA)
vector_with_na
# Removes missing values and returns rest values where length of "remove_na"<="vector_with_na"
remove_na=vector_with_na[!is.na(vector_with_na)]
remove_na
# Removes NA values as well as 0s from the vector
remove_naa=vector_with_na[!is.na(vector_with_na) & vector_with_na>0]
remove_naa
```

