---
title: "R Notebook"
output: html_notebook
---
## **Install Tidyverse, Caret, RPart for Recursive Partioning Decision Trees**
```{r setup='false'}
library(tidyverse)
library(caret)
library(rpart)
```

## **Load iris**
```{r}
df=iris
df
```
## **Check the data**
```{r}
sum(is.na(df))
sum(duplicated(df))
```
## **Rpart(Recursive partitioning is a statistical method for multivariable analysis)** is a powerful machine learning library in R that is used for building classification and regression trees. This library implements recursive partitioning and is very easy to use
```{r}
# species-dependent ad '.' is for all the columns in iris dataset, 'class' = classification, 'anova' = regression
names(df)
model=rpart(Species ~., df, method='class')
par(xpd=NA) # print tree neatly according to machine
plot(model)
text(model,col='red')

```
## **Printing model - Gives roots(alike if...else)**
```{r}
print(model)
```
## Predicts the classification value based upon thw values included from the decision tree model
```{r}
library(magrittr)
dff=data.frame(
  Sepal.Length=15.5, Sepal.Width = 3.0,
  Petal.Length = 5.2, Petal.Width = 2.0 # check decision tree for pletalength>2.45 and petalwidth>1.75
)
model %>% predict(dff,"class") # piping and predicting
```
## Better viz for decision trees
```{r}
library(rpart.plot)
rpart.plot(model)
```

